\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{\protect Acknowledgements}%



I would first like to thank my supervisor Prof.~Stephan Günnemann. I feel incredibly lucky for the opportunity to pursue a Ph.D. under your guidance. Thank you for showing me how to become a better researcher, teacher, presenter, and supervisor. I really enjoyed our brainstorming sessions where you always provided helpful and precised feedback. Thank you for your understanding and great support which was always motivating to achieve great work.

I am very grateful to all DAML group. I could not have imagined a better environment for my Ph.D. study. I would like to thank everyone in the group for their readiness to help, discuss ideas, and share research progress. This has been a lot of fun at work and during free time. I would like to particularly thank all my colleagues collaborators, and co-authors that I had chance to work with at the Technical University of Munich. Thank you Simon Geisler, Anna-Kathrin Kopetzki, Tom Wollshläger, and Nicholas Gao. Thank you also Marin Biloš for your support, help, and your great sense of humor. It was a lot of fun to share our office together. Special thank to Daniel Zügner for all your support and openness. Beyond all the great research collaborations, I would like to thank you for all the great discussions and advices.

While pursuing my Ph.D., I had the chance to work with bright, curious, and motivated students. Thank you Oliver, Maximilian, Daniel, Johannes, Sascha, Igor, Franz, Raphael, Niklas, Stephan, Simon, Sven, Chenxiang, Xun, Morgane, and John.

I am very grateful to Prof. Mykel Kochenderfer from Stanford University for hosting me during my research stay. I would like to thank everyone from SISL lab with a special thank to Ransalu Senanayake. You always make me feel welcome and provided me great advice during my stay.

I am also very grateful to the members of the Twitter Cortex team for your warm welcome and support despite the intense time at the company. I would like to particularly thank Nils Hammerla, Emanuele Rossi, Francesco Di Giovianni, and Prof. Michael Bronstein for the great discussions and research collaborations.

I would also like to express my very profound gratitude to my family, my friends, and my partner for providing me continuous encouragement, joy, and happiness through-out these years. I am particularly thankful to my parents who provided me unconditional support. I love you.

This thesis is also dedicated to my beloved grandfather Alain, godfather Bertrand, and cousin Quentin who passed away. Thank you for all the time that I could share with you. You are always a great source of inspiration for me.

%Thomas Bonald at Télé-com ParisTech who guided me during my master thesis project and allthe researchers from the LINCS who welcomed me.I  would  also  like  to  thank  my  thesis  advisors  Pawel  Herman  ofthe CSC department at KTH and Sylvain Bouveret at Ensimag whoaccompanied during my master thesis, read it carefully and gave meprecious comments on my work.I would also like to thank my examiners Johan Håstad at KTH andStephanie Hahmann at Ensimag who were involved in the validationsurvey for this research project.I must also express my very profound gratitude to my parents andto my friends for providing me continuous encouragement through-out  my  years  of  study  and  through  the  process  of  researching  andwriting this thesis.Finally, this Master’s thesis has been very interesting for many rea-sons.  First of all, I exchanged with different researchers and PhD stu-dents from many different domains.  Thus, it allows to learn a lot andimagine new ideas crossing disciplines and the results I obtained gaveme the feeling to improve the state of the art in some way.  Finally, Iam particularly happy to have submitted a paper and started collabo-rations with other researchers. For all these reasons, I plan to continuein that domain by doing a PhD.