\section{Experiments}
\label{sec:experiments_010}

We evaluate our models on large-scale synthetic and real world data. We compare to neural point process models: \textbf{RMTPP} \cite{RMTPP} and \textbf{Neural hawkes process} \cite{hawkes}. Additionally, we use various RNN models with the knowledge of the time of the next event. We measure the accuracy of class prediction, accuracy of time prediction, and evaluate on an anomaly detection task to show prediction uncertainty.

We split the data into train, validation and test set (60\%--20\%--20\%) and tune all models on a validation set using grid search over learning rate, hidden state dimension and $L_2$ regularization. After running models on all datasets $5$ times we report mean and standard deviation of test set accuracy. Details on model selection can be found in \cref{model-selection}. The code and further supplementary material is available online.\footnote{\url{https://www.daml.in.tum.de/uncertainty-event-prediction}}

We use the following data (more details in \cref{datasets}): (1) \textbf{Graph.} We generate data from a directed Erdős–Rényi graph where nodes represent the states and edges the weighted transitions between them. The time it takes to cross one edge is modelled with one normal distribution per edge. By randomly walking along this graph we created $10$K asynchronous events with $10$ unique classes.
(2) \textbf{Stack Exchange.}\footnote{\url{https://archive.org/details/stackexchange}} Sequences contain rewards as events that users get for participation on a question answering website. After preprocessing according to \cite{RMTPP} we have 40 classes and over 480K events spread over 2 years of activity of around 6700 users. The goal is to predict the next reward a user will receive.
(3) \textbf{Smart Home \normalfont\cite{SmartHome}.}\footnote{\url{https://sites.google.com/site/tim0306/datasets}} We use a recorded sequence from a smart house with 14 classes and over $1000$ events. Events correspond to the usage of different appliances. The next event will depend on the time of the day, history of usage and other appliances.
(4) \textbf{Car Indicators.} We obtained a sequence of events from car's indicators that has around $4000$ events with 12 unique classes. The sequence is highly asynchronous, with $\DeltaTime$ ranging from milliseconds to minutes.

\input{sections/010_neurips2019/paper/visualization}\label{visualization}

\input{sections/010_neurips2019/paper/event_prediction}\label{event_prediction}

\input{sections/010_neurips2019/paper/time_prediction}\label{time_prediction}

\input{sections/010_neurips2019/paper/anomaly_detection}\label{anomaly_detection}
