\textbf{\TimeScore evaluation.} Next, we aim to assess the quality of the time intervals at which we have confidence in one class. Even though \GPModel and the \DirModel do not model a distribution on time, they still have intervals at which we are certain in a class prediction, making the conditional probability a good indicator of the time occurrence of the event.

\textit{Set-up.}  While models predicting a \textit{single} time $\smash{\hat \DeltaTime_i}$ for the next event often use the MSE score $\smash{\frac{1}{n} \sum_{i=1}^n (\hat \DeltaTime_i -\DeltaTime_i^*)^2}$, in our case the MSE is not suitable since one event can occur at multiple time points. In the conventional least-squares approach, the mean of the true distribution is an optimal prediction; however, here it is almost always wrong. Therefore, we use another metric which is better suited for multimodal distributions. Assume that a model returns a score function $\smash{g_\IndexEvent^{(\IndexClass)}(\DeltaTime)}$ for each class regarding the next event $i$, where a large value means the class $\IndexClass$ is likely to occur at time $\DeltaTime$. We define $\smash{\text{\TimeScore} = \frac{1}{n} \sum_{i=1}^n \int \mathbb{1}_{g_\IndexEvent^{(\IndexClass)}(\DeltaTime) \geq g_\IndexEvent^{(\IndexClass)}(\DeltaTime_{\IndexEvent}^*)} d\DeltaTime}$. The
\TimeScore computes the size of the time intervals where the predicted score is larger than the score of the observed time $\DeltaTime_i^*$. Hence, a performant model would achieve a low \TimeScore if its score function $\smash{g_\IndexEvent^{(\IndexClass)}(\DeltaTime)}$ is high at time $\DeltaTime^*$. As the score function in our models, we use the corresponding class probability $\bar{p}_{\IndexEvent \IndexClass}(\DeltaTime)$.

\textit{Results.} We can see that our models clearly obtain the best results on all datasets. The point process version of \DirModel does not improve the performance. Thus, taking also into account the class prediction performance, we recommend to use our other two models. In Appendix \ref{time_mse} we compare FD-Dir-PP with other neural point process models on  time prediction using the MSE score and achieve similar results.
% \input{time_score}